Proof equivalence in \capsabbrev{mll} is \capsabbrev{pspace}-complete

### 2015-07-08 Reviews LMCS/Response.txt

We thank the referees for their careful reading of our paper, and for the detailed and helpful feedback they have provided. Here, we will discuss each review in turn, as well as the changes we have made in response. We omit the suggested minor changes that we have implemented, but comment on those we chose not to.

@@

*** We have expanded the discussion of these alternatives, giving brief hints as to how they avoid the unit problem.

I am also surprised by the total absence of a discussion of the case of MLL with the MIX rules.

*** We opted not to discuss MIX - depending on which rules are added, either the units are equated, which is semantically equivalent to omitting them, or they are not, in which case the logic is conservative over MLL and the unit problem remains.

1) Numerical constraints [...]

*** We have added a further illustration showing where the multiplication by 3x the number of edges is needed.

2) Parity [...]

*** We have added some introductory paragraphs to the subsection on parity, and adjusted Definition 37 for clarity (see also the response to Review 1 above).

3) Soundness/readback[...]

*** We have put the read-back or decoding function in a separate proposition, with justification. The only concern for decoding is the following lemma: if two linkings are equivalent, there is a reconfiguration path between their decodings.

First, the authors choose to restrict jumps to target 1-­formulae and to rewire jumps according to a big-­step strategy, building on the work of Hughes. [...]

*** We now have more clearly motivated this choice: it is mostly cosmetic, as it simplifies the diagrammatic notation for proof nets - though it does impact the formulation of some lemmas in the section "Rewiring proof nets".

Second, in Section 4 it is said that the reconfiguration problem remains PSPACE-­complete under a number of restrictions, e.g. fixing constraint at the value 2, but the authors choose the encode the general case, without restrictions. [...]

*** We don't think these restrictions would actually simplify the encoding - in fact, it would be exactly the same. It would still be necessary, for instance, to correct for parity, and to multiply weights by 3 times the number of edges.

### 2015-07-08 main/LMCS.tex
final?

Below top is a weight element, below bottom a constraint element.
\[
\begin{array}{ccc}
	  \vc{\begin{tikzpicture}[x=5mm,y=5mm,octo]
	 	\node[big_] (A) at (0,0) {$i\ap$}; 
	 	\node[big_] (B) at (2,0) {$j\ap$};
		\node[big_] (C) at (4,0) {$k\ap$};
		\draw[rounded corners] (-1,-1) rectangle (5,1);
	  \end{tikzpicture}}
	&=&
	  \bigtn(\bot^{i+1})~~\pr~~\bigtn(\bot^{j+1})~~\pr~~\bigtn(\bot^{k+1})
\\ \\
	  \vc{\begin{tikzpicture}[x=5mm,y=5mm,octo]
		\node[big1] (1) at (0,0) {$m\ap$}; 
		\node[big1] (2) at (3,0) {$n\ap$};
		\draw[rounded corners] (-1,-1) rectangle (1,1) (2,-1) rectangle (4,1);
		\draw (1,0)--(2,0);
	  \end{tikzpicture}}
	&=&
	  \bigparr(\1^{m+1})~\tn~\bigparr(\1^{n+1})
\end{array}
\]
For all edges and vertices in the encoding of a constraint graph, the sum $i+j+k=m+n$ will be the same---this way, a priori any weight element may connect to any constraint element. The value $m$ will be unique for each vertex, and used for all constraint elements in the encoding of that vertex. The weight element above is for an edge connecting the two vertices for which $m=i$ and $m=i+j$: with those constraint elements it is possible to form a proof net, as illustrated below.
\[
	\connectingElements 1 \qquad \connectingElements 2
\]
To ensure that a weight element may not form a proof net with a constraint element of another vertex, the values of $m$ and $n$ are chosen such that $m\equiv1$ and $n\equiv2 \pmod 3$, and accordingly $i\equiv1$, $j\equiv0$, and $k\equiv2\pmod 3$. It then becomes impossible for $m$ to equal any other sum over $i$, $j$, and $k$, so that forming a proof net requires either $m=i$ or $m=i+j$.

@@

An edge in a constraint graph will be encoded by an \emph{edge-gadget}, illustrated below left, constructed by stringing together a number of similar weight elements, plus a single $\bot$-formula which we will refer to as the \emph{indicator}. Illustrated below right is a \emph{vertex-gadget} encoding a vertex, formed by a number of constraint elements plus a single \emph{indicator target}.
\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[big_] (A) at (0,0) {$i\ap$}; 
	\node[big_] (B) at (2,0) {$j\ap$};
	\node[big_] (C) at (4,0) {$k\ap$};
	\draw[rounded corners] (-1,-1) rectangle (5,1);
	\node[big_] (D) at (0,4) {$i\ap$}; 
	\node[big_] (E) at (2,4) {$j\ap$};
	\node[big_] (F) at (4,4) {$k\ap$};
	\draw[rounded corners] (-1,3) rectangle (5,5);
	\draw (2,1)--(2,1.5) (2,2.5)--(2,3);
	\draw[dotted] (2,1.5)--(2,2.5);
	\draw (2,-1)--(2,-2) node[bullet] {};
	\path (0,-2.5)--(1,-2.5);
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[big1] (1) at (0,0) {$m\ap$}; 
	\node[big1] (2) at (3,0) {$n\ap$};
	\draw[rounded corners] (-1,-1) rectangle (1,1) (2,-1) rectangle (4,1);
	\draw (1,0)--(2,0);
	\node[big1] (3) at (0,4) {$m\ap$};
	\node[big1] (4) at (3,4) {$n\ap$};
	\draw[rounded corners] (-1,3) rectangle (1,5) (2,3) rectangle (4,5);
	\draw (1,4)--(2,4);
	\draw[dotted] (1.5,1.5)--(1.5,2.6);
	\node[circ] at (1.5,-2) {};
	\draw[rounded corners] (-1.5,-2.5) rectangle (4.5,5.5);
\end{tikzpicture}
\]
The natural way for edge-gadgets to link up with a vertex-gadget is shown in Figure~\ref{fig:connect example}: indicators connect to indicator targets, and weight elements connect to constraint elements or weight absorbers. The illustration depicts two equivalent proof nets, both encoding the same (fragment of a) constraint graph: a single vertex that is the target of a weight-1 edge (on the left, in red) and a weight-2 edge (on the right, in blue). The inflow of the vertex is 3, and its constraint is 2; then 2 weight elements are connected to the constraint elements in the vertex-gadget, while one remains free (connected to weight absorbers). It is essential to the functioning of the encoding that any re-distribution of weight elements gives an equivalent proof net, as illustrated by the example. Then in the second proof net, the single weight element of the left edge-gadget is free, which means the indicator jump can be rewired to connect to a different vertex.

@@

In an inappropriate linking as shown above, since both halves of the constraint element are connected, the weight elements must be disconnected---otherwise, the switching condition would be violated. That means the weight elements must belong to different edges.

As the linkings above illustrate, it may occur that one subformula of a weight element $A\pr B\pr C$ fills one half of a constraint element. Consequently, a weight element can fill three halves, but only of different constraint elements. The other three halves may be filled by weight elements of a different edge---so to fill 3 constraint elements requires 2 inappropriate edges. To fill the next three constraint elements, at most 1 previous inappropriate edge may be used, and one additional one is needed. To fill $3n$ constraint elements inappropriately therefore requires $n+1$ edges. It thus suffices to encode one unit of weight on an edge by $3\times|E|$ weight elements (where $|E|$ is the number of edges in the constraint graph), and correspondingly to encode one unit of inflow constraint by $3\times|E|$ constraint elements.

@@

The complete encoding of a constraint graph $G$ will then be a sequent $\Gamma$ consisting of:
%
\begin{enumerate}
 	\item
all vertex-gadgets, combined in a single formula via tensors, 
	\item
all edge-gadgets as individual formulae, and
	\item
a sufficient number of weight absorbers ($\1$-formulae).
\end{enumerate}
%
A configuration for $G$ will be encoded as a proof net for $\Gamma$, and conversely each proof net for $\Gamma$ may be interpreted as a (partial) configuration for $G$. The encoding will be made formal in the next section. %Section~\ref{sec:correctness}.

@@

We will formalise the encoding of \capsabbrev{ncg}-reconfiguration into \capsabbrev{mll} proof equivalence that was informally introduced in the previous section. %Section~\ref{sec:encoding}.
A constraint graph $G$ will be encoded as a sequent $\itn G$, and a configuration $\gamma$ for $G$ will be encoded as a proof net $\itn\gamma$ for $\itn G$. We will show that $\gamma\perm\delta$ if and only if $\itn\gamma\perm\itn\delta$ (modulo a small adjustment to ensure even parity between $\itn\gamma$ and $\itn\delta$).

@@

\noindent
The encoding of a vertex $v_k$ is the formula
\[
	\itn{v_k} = \bigparr\big(C_n(k)^{3m\,\times\,c(v_k)} \big)\pr\1
\]
where each \emph{constraint element} $C_n(k)$ is the formula
\[
	C_n(k) = \bigparr\big(1^{3k+2}\big) \tn \bigparr\big(1^{3(n-k)+3}\big)~.
\]

\noindent
The encoding of an edge $e$ connecting vertices $v_i$ and $v_j$ with $i\leq j$ is the formula
\[
	\itn{e} = \bigtn\big(W_n(i,j)^{3m\,\times\,w(e)}\big)\tn\bot
\]
where each \emph{weight element} $W_n(i,j)$ is the formula
\[
	W_n(i,j) = \bigtn\big(\bot\!^{3i+2}\big)\pr\bigtn\big(\bot\!^{3(j-i)+1}\big)\pr\bigtn\big(\bot\!^{3(n-j)+3}\big)\,.
\]

@@

\begin{definition}
\label{def:standard linkings}
For $W=W_n(i,j)=A\pr B\pr C$ a weight element,
\begin{enumerate}
	\item
for a constraint element $V=C_n(i)=X\tn Y$, the \emph{standard linking} for the sequent $W,V$ links the first $\bot$ in $A$ to the first $\1$ in $X$, the first $\bot$ in $B$ and $C$ each to the first $\1$ in $Y$, and each remaining $\bot$ in $A,B,C$ to a remaining $\1$ in $X,Y$ in their order of occurrence;
	\item
for $V=C_n(j)=X\tn Y$, the \emph{standard linking} for $W,V$ is defined as above, except the first $\bot$ in $B$ links to the first $\1$ in $X$;
	\item
the \emph{standard linking} for the sequent $W,\1^{3n+4}$ links the first $\bot$ in $A$, $B$, and $C$ to the first $\1$, and each remaining $\bot$ to a remaining $\1$ in order of occurrence.
\end{enumerate}
\end{definition}

@@

\begin{definition}
\label{def:configuration encoding}
The \emph{encoding} $\itn\gamma$ of a total configuration $\gamma$ for a constraint graph $G$ is a linking $\links$ for $\itn G$, constructed incrementally for each successive edge $e$ and for each successive weight element $W$ within $e$, as follows.
%
Let $\gamma(e)=v$; firstly, the indicator of $\itn e$ is linked to the indicator target of $\itn v$.
%
Then successively for each weight element $W$ in $e$, if $\itn v$ has a first free constraint element $C$, extend $\itn\gamma$ to include the standard linking on $W,C$; otherwise, extend $\itn\gamma$ by the standard linking on the sequent consisting of $W$ plus the first $3n+4$ free weight absorbers.
\end{definition}

@@

\begin{proof}
Using Proposition~\ref{prop:scoping correctness}, it is sufficient to give a suitable box for each $\pr$. 
%
The box of each weight element $W$ is the sequent $W,C$ or $W,1^{3n+4}$ of its standard linking, which forms a proof net by Proposition~\ref{prop:element linkings}.
%
The box of each vertex-gadget $\itn v$ contains the edge-gadgets $\itn e$ such that $\gamma(e)=v$, plus all the weight absorbers within boxes of weight elements inside $\itn e$.
%
Since the weights of the connected edges $e$ sum to more than the inflow constraint of $v$, there are no unused constraint elements remaining in $\itn v$.
%
After closing the box of each $W$, each edge-gadget in the box of $\itn v$ becomes a single string of connected vertices, connected to other edge-gadgets only via the indicator target of $\itn v$, thus forming a tree.
\end{proof}

Finally, we will encode an instance of the NCL reconfiguration problem: a pair of configurations $\gamma$ and $\delta$ for a given NCL graph $G$, for which the problem asks whether $\gamma\perm\delta$. Here we need to ensure that the two encodings as proof nets have even parity; otherwise, they will never be equivalent. We have opted for a simple and straightforward solution: first we encode both configurations as $\itn\gamma$ and $\itn\delta$; then we check their parity, and if it is odd, we adjust it by swapping two weight absorbers on $\itn\delta$.

%Finally, if two connected configurations $\gamma\perm\delta$ are encoded individually as $\itn\gamma$ and $\itn\delta$, these may not be equivalent simply because their parity happens to be odd. For that reason we shall adjust the encoding of the second to guarantee an even parity.

\begin{definition}
%
The \emph{encoding} of an instance of NCL-reconfiguration $(G,\gamma,\delta)$ is the triple $(\itn G,\itn\gamma,\itn\delta_\gamma)$ where
\[
	\itn\delta_\gamma=\left\{
	 \begin{array}{ll}\itn\delta & \text{if $\itn\gamma$ and $\itn\delta$ have even parity} \\
	 				  \itn\delta' & \text{otherwise}\end{array}\right.
\]
and $\itn\delta'$ is $\itn\delta$ with the first two weight absorbers exchanged.
\end{definition}

%Let $\itn\delta'$ be $\itn\delta$ with the first two weight absorbers exchanged. For configurations $\gamma$ and $\delta$ for $G$, let $\itn\delta_\gamma$ be $\itn\delta$ if it has even parity with $\itn\gamma$, and $\itn\delta'$ otherwise.
%

In the remainder, we will show that our encoding is correct, i.e.\ that $\gamma\perm\delta$ if and only if $\itn\gamma\perm\itn\delta_\gamma$.
%
This will be separated into two parts: completeness ($\Rightarrow$) and soundness ($\Leftarrow$) .

@@

\begin{lemma}
\label{lem:octopus roll}
Let $\links$ be a proof net for $\Gamma=\itn v,\itn{e_1},\dotsc,\itn{e_m},1^p$ such that each edge-gadget $\itn e_i$ is well linked. Let $W_i$ and $W_j$ be weight elements in edge-gadgets $\itn{e_i}$ and $\itn{e_j}$ respectively; let $W_i$ be linked to $V$ in $\itn v$, and let $W_j$ be linked to weight absorbers $1^n$. Then there is a net $\links'\perm\links$ in which $W_j$ is well linked to $V$, $W_i$ is linked to $1^n$, and $\links'$ agrees with $\links$ otherwise.
\end{lemma}

@@

\begin{proof}
Let $W_i=A\pr B\pr C$, $W_j=D\pr E\pr F$, and $V=X\tn Y$.
%
The rewiring path will be illustrated for the case where $A,B,X$ and $D,X$, and thus also $C,Y$ and $E,F,Y$, are balanced sequents; other cases are similar.

@@

\noindent
\textbf{2.}\quad The jump $\jump iv$ is rewired to a weight absorber together with only the jumps from the first $\bot$ of each of $D$, $E$, and $F$:

@@

\noindent
\textbf{3.}\quad The jump from the first $\bot$ of $D$ is moved to $X$, and those from the first $\bot$ of $E$ and $F$ are moved to $Y$:

@@

\noindent
\textbf{4.}\quad In the present configuration, the jumps from the weight elements form two subnets: one over the sub-sequent $A,B,X,D,1^m$, and one over the sub-sequent $C,Y,E,F,1^k$, for some $m$ and $k$. By Proposition~\ref{prop:parity determines equivalence}, these subnets are equivalent to any other over the same sequent, as long as their parity is preserved. It is then sufficient to choose linkings so that $D\pr E\pr F$ is in a standard linking with $X\tn Y$, with a minor adjustment: two jumps from $D$ to $X$ should remain exchanged, compared to the standard linking, for step 6 below. The jumps of $A,B,C$ link to the weight absorbers $1^m$ and $1^k$, with one remaining jump from $B$ to $X$ and one from $C$ to $Y$.

@@

\noindent
\textbf{5.}\quad The jump from $C$ to $Y$ is moved towards a weight absorber connected to $A,B$.

@@

\noindent
\textbf{6.}\quad The jump from $B$ to $X$ is the one remaining connection between the edge-gadgets $\itn{e_i}$ and $\itn{e_j}$. Lemma~\ref{lem:double exchange} allows to swap the targets of the jump from $B$ to $X$ and the jump from $i$, and simultaneously undo the exchange in the jumps from $D$ to $X$ added in step 4 above.

@@

It will be shown that proof net rewiring $\itn\gamma\perm\itn\delta_\gamma$ is sound for \capsabbrev{ncg}-reconfiguration, $\gamma\perm\delta$.
%
To each proof net $\links$ for an encoded constraint graph $\itn G$ we will associate a partial configuration $\gamma=\coitn\links$, such that 1) a rewiring step between $\links$ and $\links'$ corresponds to a reconfiguration path $\coitn\links\perm\coitn{\links'}$, and 2) the function $\coitn-$ is a retraction of the encoding of configurations, $\coitn{\itn\gamma}=\gamma$.

The configuration $\coitn\links$ will assign an edge $e$ to a vertex $v$ when, in the proof net $(\links,\itn G)$, the edge-gadget $\itn e$ is in the empire of the vertex-gadget $\itn v$.
%
Firstly, it will be shown that $\coitn\links$ is a partial function.

>>>>

Below top is a weight element, below bottom a constraint element.

\[
\begin{array}{ccc}
	  \vc{\begin{tikzpicture}[x=5mm,y=5mm,octo]
	 	\node[big_] (A) at (0,0) {$i\ap$}; 
	 	\node[big_] (B) at (2,0) {$j\ap$};
		\node[big_] (C) at (4,0) {$k\ap$};
		\draw[rounded corners] (-1,-1) rectangle (5,1);
	  \end{tikzpicture}}
	&=&
	  \bigtn(\bot^{i+1})~~\pr~~\bigtn(\bot^{j+1})~~\pr~~\bigtn(\bot^{k+1})
\\ \\
	  \vc{\begin{tikzpicture}[x=5mm,y=5mm,octo]
		\node[big1] (1) at (0,0) {$m\ap$}; 
		\node[big1] (2) at (3,0) {$n\ap$};
		\draw[rounded corners] (-1,-1) rectangle (1,1) (2,-1) rectangle (4,1);
		\draw (1,0)--(2,0);
	  \end{tikzpicture}}
	&=&
	  \bigparr(\1^{m+1})~\tn~\bigparr(\1^{n+1})
\end{array}
\]

\noindent
For all edges and vertices in the encoding of a constraint graph, the sum $i+j+k=m+n$ will be the same---this way, a priori any weight element may connect to any constraint element. The value $m$ will be unique for each vertex, and used for all constraint elements in the encoding of that vertex. The weight element above is for an edge connecting the two vertices for which $m=i$ and $m=i+j$: with those constraint elements it is possible to form a proof net, as illustrated below.

\[
	\connectingElements 1 \qquad \connectingElements 2
\]

\noindent
To ensure that a weight element may not form a proof net with a constraint element of another vertex, the values of $m$ and $n$ are chosen such that $m\equiv1$ and $n\equiv2 \pmod 3$, and accordingly $i\equiv1$, $j\equiv0$, and $k\equiv2\pmod 3$. It then becomes impossible for $m$ to equal any other sum over $i$, $j$, and $k$, so that forming a proof net requires either $m=i$ or $m=i+j$.

@@

An edge in a constraint graph will be encoded by an \emph{edge-gadget}, illustrated below left, constructed by stringing together a number of similar weight elements, plus a single $\bot$-formula which we will refer to as the \emph{indicator}. Illustrated below right is a \emph{vertex-gadget} encoding a vertex, formed by a number of constraint elements plus a single \emph{indicator target}.

\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[big_] (A) at (0,0) {$i\ap$}; 
	\node[big_] (B) at (2,0) {$j\ap$};
	\node[big_] (C) at (4,0) {$k\ap$};
	\draw[rounded corners] (-1,-1) rectangle (5,1);
	\node[big_] (D) at (0,4) {$i\ap$}; 
	\node[big_] (E) at (2,4) {$j\ap$};
	\node[big_] (F) at (4,4) {$k\ap$};
	\draw[rounded corners] (-1,3) rectangle (5,5);
	\draw (2,1)--(2,1.5) (2,2.5)--(2,3);
	\draw[dotted] (2,1.5)--(2,2.5);
	\draw (2,-1)--(2,-2) node[bullet] {};
	\path (0,-2.5)--(1,-2.5);
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[big1] (1) at (0,0) {$m\ap$}; 
	\node[big1] (2) at (3,0) {$n\ap$};
	\draw[rounded corners] (-1,-1) rectangle (1,1) (2,-1) rectangle (4,1);
	\draw (1,0)--(2,0);
	\node[big1] (3) at (0,4) {$m\ap$};
	\node[big1] (4) at (3,4) {$n\ap$};
	\draw[rounded corners] (-1,3) rectangle (1,5) (2,3) rectangle (4,5);
	\draw (1,4)--(2,4);
	\draw[dotted] (1.5,1.5)--(1.5,2.6);
	\node[circ] at (1.5,-2) {};
	\draw[rounded corners] (-1.5,-2.5) rectangle (4.5,5.5);
\end{tikzpicture}
\]

The natural way for edge-gadgets to link up with a vertex-gadget is shown in Figure~\ref{fig:connect example}: indicators connect to indicator targets, and weight elements connect to constraint elements or weight absorbers. The illustration depicts two equivalent proof nets, both encoding the same (fragment of a) constraint graph: a single vertex that is the target of a weight-1 edge (on the left, in red) and a weight-2 edge (on the right, in blue). The inflow of the vertex is 3, and its constraint is 2; then 2 weight elements are connected to the constraint elements in the vertex-gadget, while one remains free (connected to weight absorbers). It is essential to the functioning of the encoding that any re-distribution of weight elements gives an equivalent proof net, as illustrated by the example. Then in the second proof net, the single weight element of the left edge-gadget is free, which means the indicator jump can be rewired to connect to a different vertex.

@@

In an inappropriate linking as shown above, since both halves of the constraint element are connected, the weight elements must be disconnected---otherwise, the switching condition would be violated. That means the weight elements must belong to different edges. Note that it is possible to use more than one weight element from the same edge to fill one half of a constraint element; or even weight elements from different edges, but we are interested here only in the \emph{minimum} number of edges needed to inappropriately fill a vertex-gadget.

%Since the different halves of a constraint element must be filled by different edge-gadgets, the switching condition forces inappropriately connected edge-gadgets and constraint elements to form a tree, with edge-gadgets as nodes and constraint elements as edges. Since in a tree the number of nodes is always one greater than the number of edges, it requires $n+1$ edge-gadgets to fill $n$ constraint elements. It thus suffices to multiply the number of weight elements by the number of edges: we will encode one unit of weight on an edge by $|E|$ weight elements (where $|E|$ is the number of edges in the constraint graph), and correspondingly encode one unit of inflow constraint by $|E|$ constraint elements.

As the linkings above illustrate, it may occur that one subformula of a weight element $A\pr B\pr C$ fills one half of a constraint element. In the pathological case for $(v_i)$ where $i$ is very small or very large, a weight element can fill three halves of different constraint elements, as illustrated in Figure~\ref{fig:pathological}. As in the illustration, the other three halves may be filled by weight elements of a different edge---so to fill 3 constraint elements requires 2 inappropriate edges. To fill the next three constraint elements, at most 1 previous inappropriate edge may be used, and one additional one is needed. To fill $3n$ constraint elements inappropriately therefore requires $n+1$ edges. It thus suffices to encode one unit of weight on an edge by $3\times|E|$ weight elements (where $|E|$ is the number of edges in the constraint graph), and correspondingly to encode one unit of inflow constraint by $3\times|E|$ constraint elements.

\begin{figure}
\[
\begin{tikzpicture}[x=-5mm,y=-5mm,octo]
	\node[big1] (1) at (0,3) {};
	\foreach \i in {0,3,6} {
		\node[big_] (A\i) at (3.5,\i) {}; 
		\node[big_] (B\i) at (5,  \i) {}; 
		\node[big_] (C\i) at (7,  \i) {};
		\node[big_] (D\i) at (9,  \i) {};
		\draw (A\i)--(B\i);
		\draw[rounded corners] ($(2.5,-1) + (0,\i)$) rectangle ($(10,1) + (0,\i)$);
		%
		\node[big1] (2\i) at (13,\i) {}; 
		\node[big1] (3\i) at (16,\i) {};
		\draw[rounded corners] ($(12,-1) + (0,\i)$) rectangle ($(14,1) + (0,\i)$);
		\draw[rounded corners] ($(15,-1) + (0,\i)$) rectangle ($(17,1) + (0,\i)$);
		\draw (14,\i)--(15,\i);
		%
		\begin{scope}[thepink,->,big>]
			\draw[bend left=30] (B\i) to (2\i);
			\draw[bend left=15]  (C\i) to (2\i);
			\draw (D\i) to (2\i);
		\end{scope}
	}
	\begin{scope}[thepink,->,big>]
		\draw[bend right=20] (A6) to (1);
		\draw                (A3) -- (1);
		\draw[bend left=20]  (A0) to (1);
	\end{scope}
	\draw (6.25,1) -- +(0,1) (6.25,4) -- +(0,1);
	\draw[rounded corners] (11.5,-1.5) rectangle (17.5,7.5);
	\foreach \i in {1,3,5} {
		\node[big_] (E\i) at (20,  \i) {};
		\node[big_] (F\i) at (21.5,\i) {};
		\draw (E\i)--(F\i);
	}
	\draw[rounded corners] (19,0) rectangle (22.5,6);
	\node[big1] (4) at (24,3) {};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=20] (E5) to (36);
		\draw                (E3) -- (33);
		\draw[bend left=20]  (E1) to (30);
		\draw[bend left=20]  (F5) to (4);
		\draw                (F3) -- (4);
		\draw[bend right=20] (F1) to (4);
	\end{scope}
\end{tikzpicture}
\]
\caption{A weight element connecting to three constraint elements}
\label{fig:pathological}
\end{figure}

@@

The complete encoding of a constraint graph $G$ will then be a sequent $\Gamma$ consisting of:
%
\begin{enumerate}
 	\item
all vertex-gadgets, combined in a single formula via tensors, 
	\item
all edge-gadgets as individual formulae, and
	\item
a sufficient number of weight absorbers ($\1$-formulae).
\end{enumerate}
%
A configuration for $G$ will be encoded as a proof net for $\Gamma$, and conversely each proof net for $\Gamma$ may be interpreted as a (partial) configuration for $G$. The encoding will be made formal in the next section.

@@

We will formalise the encoding of \capsabbrev{ncg}-reconfiguration into \capsabbrev{mll} proof equivalence that was informally introduced in the previous section. A constraint graph $G$ will be encoded as a sequent $\itn G$, and a configuration $\gamma$ for $G$ will be encoded as a proof net $\itn\gamma$ for $\itn G$. We will show that $\gamma\perm\delta$ if and only if $\itn\gamma\perm\itn\delta$ (modulo a small adjustment to ensure even parity between $\itn\gamma$ and $\itn\delta$).

@@

\noindent
The encoding of a vertex $v_k$ is the formula
\[
	\itn{v_k} = \bigparr\big(C(k,n)^{3m\,\times\,c(v_k)} \big)\pr\1
\]
where each \emph{constraint element} $C(k,n)$ is the formula
\[
	C_n(k) = \bigparr\big(1^{3k+2}\big) \tn \bigparr\big(1^{3(n-k)+3}\big)~.
\]

\noindent
The encoding of an edge $e$ connecting vertices $v_i$ and $v_j$ with $i\leq j$ is the formula
\[
	\itn{e} = \bigtn\big(W(i,j,n)^{3m\,\times\,w(e)}\big)\tn\bot
\]
where each \emph{weight element} $W(i,j,n)$ is the formula
\[
	W(i,j,n) = \bigtn\big(\bot\!^{3i+2}\big)\pr\bigtn\big(\bot\!^{3(j-i)+1}\big)\pr\bigtn\big(\bot\!^{3(n-j)+3}\big)\,.
\]

@@

\begin{definition}
\label{def:standard linkings}
For $W=W(i,j,n)=X\pr Y\pr Z$ a weight element,
\begin{enumerate}
	\item
for a constraint element $C=C(i,n)=A\tn B$, the \emph{standard linking} for the sequent $C,W$ links the first $\bot$ in $X$ to the first $\1$ in $A$, the first $\bot$ in $Y$ and $Z$ each to the first $\1$ in $B$, and each remaining $\bot$ in $X,Y,Z$ to a remaining $\1$ in $A,B$ in their order of occurrence;
	\item
for $C=C(j,n)=A\tn B$, the \emph{standard linking} for $C,W$ is defined as above, except the first $\bot$ in $Y$ links to the first $\1$ in $A$;
	\item
the \emph{standard linking} for the sequent $W,\1^{3n+4}$ links the first $\bot$ in $X$, $Y$, and $Z$ to the first $\1$, and each remaining $\bot$ to a remaining $\1$ in order of occurrence.
\end{enumerate}
\end{definition}

@@

\begin{definition}
\label{def:configuration encoding}
The \emph{encoding} $\itn\gamma$ of a total configuration $\gamma$ for a constraint graph $G$ is a linking $\links$ for $\itn G$, constructed incrementally for each successive edge $e$ and for each successive weight element $W$ within $e$, as follows.
%
Let $\gamma(e)=v$; firstly, the indicator of $\itn e$ is linked to the indicator target of $\itn v$.
%
Then successively for each weight element $W$ in $e$, if $\itn v$ has a first free constraint element $C$, extend $\itn\gamma$ to include the standard linking on $C,W$; otherwise, extend $\itn\gamma$ by the standard linking on the sequent consisting of $W$ plus the first $3n+4$ free weight absorbers.
\end{definition}

@@

\begin{proof}
Using Proposition~\ref{prop:scoping correctness}, it is sufficient to give a suitable box for each $\pr$. 
%
The box of each weight element $W$ is the sequent $C,W$ or $W,1^{3n+4}$ of its standard linking, which forms a proof net by Proposition~\ref{prop:element linkings}.
%
The box of each vertex-gadget $\itn v$ contains the edge-gadgets $\itn e$ such that $\gamma(e)=v$, plus all the weight absorbers within boxes of weight elements inside $\itn e$.
%
Since the weights of the connected edges $e$ sum to more than the inflow constraint of $v$, there are no unused constraint elements remaining in $\itn v$.
%
After closing the box of each $W$, each edge-gadget in the box of $\itn v$ becomes a single string of connected vertices, connected to other edge-gadgets only via the indicator target of $\itn v$, thus forming a tree.
\end{proof}

Finally, we will encode an instance of the NCL reconfiguration problem, consisting of an NCL graph $G$ with two configurations $\gamma$ and $\delta$, for which the problem asks whether $\gamma\perm\delta$. Here we need to ensure that the two encodings as proof nets have even parity; otherwise, they will never be equivalent. We have opted for a simple and straightforward solution: first we encode both configurations as $\itn\gamma$ and $\itn\delta$; then we check their parity, and if it is odd, we adjust it by swapping two weight absorbers on $\itn\delta$.

\begin{definition}
%
The \emph{encoding} of an instance of NCL-reconfiguration $(G,\gamma,\delta)$ is the triple $(\itn G,\itn\gamma,\itn\delta_\gamma)$ where
\[
	\itn\delta_\gamma=\left\{
	 \begin{array}{ll}\itn\delta & \text{if $\itn\gamma$ and $\itn\delta$ have even parity} \\
	 				  \itn\delta' & \text{otherwise}\end{array}\right.
\]
and $\itn\delta'$ is $\itn\delta$ with the first two weight absorbers $a$ and $b$ exchanged: $v\mathbin{\xjump}a$ in $\itn\delta'$ if and only if $v\mathbin{\xjump}b$ in $\itn\delta$, and vice versa.
\end{definition}

In the remainder, we will show that our encoding is correct, i.e.\ that $\gamma\perm\delta$ if and only if $\itn\gamma\perm\itn\delta_\gamma$. This will be separated into two parts: completeness ($\Rightarrow$) and soundness ($\Leftarrow$).

@@

\begin{lemma}
\label{lem:octopus roll}
Let $\links$ be a proof net for $\Gamma=\itn v,\itn{e_1},\dotsc,\itn{e_m},1^p$ such that each edge-gadget $\itn e_i$ is well linked. Let $W_i$ and $W_j$ be weight elements in edge-gadgets $\itn{e_i}$ and $\itn{e_j}$ respectively; let $W_i$ be linked to $C$ in $\itn v$, and let $W_j$ be linked to weight absorbers $1^n$. Then there is a net $\links'\perm\links$ in which $W_j$ is well linked to $C$, $W_i$ is linked to $1^n$, and $\links'$ agrees with $\links$ otherwise.
\end{lemma}

@@

\begin{proof}
Let $W_i=X\pr Y\pr Z$, $W_j=P\pr Q\pr R$, and $C=A\tn B$.
%
The rewiring path will be illustrated for the case where $X,Y,A$ and $P,A$, and thus also $Z,B$ and $Q,R,B$, are balanced sequents; other cases are similar.

@@

\noindent
\textbf{2.}\quad The jump $\jump iv$ is rewired to a weight absorber, together with only the jumps from the first $\bot$ of each of $P$, $Q$, and $R$:

@@

\noindent
\textbf{3.}\quad The jump from the first $\bot$ of $P$ is moved to $A$, and those from the first $\bot$ of $Q$ and $R$ are moved to $B$:

@@

\noindent
\textbf{4.}\quad In the present configuration, the jumps from the weight elements form two subnets: one over the sub-sequent $X,Y,A,P,1^m$, and one over the sub-sequent $Z,B,Q,R,1^k$, for some $m$ and $k$. By Proposition~\ref{prop:parity determines equivalence}, these subnets are equivalent to any other over the same sequent with which they have even parity. It is then sufficient to choose linkings so that $P\pr Q\pr R$ is in a standard linking with $A\tn B$, with a minor adjustment: two jumps from $P$ to $A$ should remain exchanged, compared to the standard linking, for step 6 below. The jumps of $X,Y,Z$ link to the weight absorbers $1^m$ and $1^k$, with one remaining jump from $Y$ to $A$ and one from $Z$ to $B$.

@@

\noindent
\textbf{5.}\quad The jump from $Z$ to $B$ is moved towards a weight absorber connected to $X,Y$.

@@

\noindent
\textbf{6.}\quad The jump from $Y$ to $A$ is the one remaining connection between the edge-gadgets $\itn{e_i}$ and $\itn{e_j}$. Lemma~\ref{lem:double exchange} allows to swap the targets of the jump from $Y$ to $A$ and the jump from $i$, and simultaneously undo the exchange in the jumps from $P$ to $A$ added in step 4 above.

@@

It will be shown that proof net rewiring $\itn\gamma\perm\itn\delta_\gamma$ is sound for \capsabbrev{ncg}-reconfiguration, i.e\ that it implies $\gamma\perm\delta$.
%
To each proof net $\links$ for an encoded constraint graph $\itn G$ we will associate a partial configuration $\gamma=\coitn\links$, such that 1) a rewiring step between $\links$ and $\links'$ corresponds to a reconfiguration path $\coitn\links\perm\coitn{\links'}$, and 2) the function $\coitn-$ is a retraction of the encoding of configurations, $\coitn{\itn\gamma}=\gamma$.

The configuration $\coitn\links$ will assign an edge $e$ to a vertex $v$ when, in the proof net $(\itn G,\links)$, the edge-gadget $\itn e$ is in the empire of the vertex-gadget $\itn v$.
%
Firstly, it will be shown that $\coitn\links$ is a partial function.
